---
title: "HW 8"
author: "Brandon Fenton and Matt Pettigrew"
date: "November 26, 2016"
output:
  pdf_document:
    fig_height: 3
    fig_width: 5
  html_document:
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5

header-includes: \usepackage{float} \usepackage{bm} \usepackage{amsmath} \usepackage{amssymb} \usepackage{microtype} \usepackage{wasysym}
---

```{r setup, include=FALSE, echo=F}
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)

library(grid)
library(gridExtra)
library(pander)
library(dplyr)
library(ggplot2)
library(effects)
library(ggfortify)
library(parallel)

panderOptions('missing', "-")

pander_lm <-function (fit, ...)
{
  fit.sum <- summary(fit)
  fit.coef <- fit.sum$coefficients
  fit.ttable <- matrix(nrow=length(fit.sum$aliased), ncol=4)
  colnames(fit.ttable) <- colnames(fit.sum$coefficients)
  rownames(fit.ttable) <- names(fit.sum$aliased)

  notna <- as.vector(which(!fit.sum$aliased))
  fit.ttable[notna,] <- fit.coef
  fit.ttable <- as.data.frame(fit.ttable)
  fit.ttable$`Pr(>|t|)` <- ifelse(fit.ttable$`Pr(>|t|)` < 0.0001, "<0.0001",
                                     sprintf("%.4f", fit.ttable$`Pr(>|t|)`))
  

  pander(fit.ttable, ...)
}

pander_anova <-function (fit, ...)
{
  fit.anova <- anova(fit)
  fit.anova$`Pr(>F)` <- ifelse(fit.anova$`Pr(>F)` < 0.0001, "<0.0001",
                                  sprintf("%.4f", fit.anova$`Pr(>F)`))

pander(fit.anova, ...)
}

clust <- makeCluster(detectCores())

```

1) _What do they provide for a reason for removing the quadratic trend? Why do they want to use a trend model with limited flexibility?_

    The authors justify removing the quadratic trend on the basis that ARMA models rely on the assumption of stationarity.  The goal in removing the trend is to obtain residuals with a constant mean of zero and constant variance.  If the residuals satisfy these properties then it stationarity is assumed, allowing an ARMA model to be utilized.  Since the goal was to remove only the trend and not the variation around the trend, "simple, parsimonious methods" were given preference.  This rules out many possible detrending methods; in particular methods with too much flexibility (for example, non-stiff splines) are unsuitable since they can lead to overfitting.
    
2) _How do they determine the order of the trend model from none to quadratic? Are there any potential criticisms of the way they do this? Do not discuss the potential issue with doing these tests without accounting for potential autocorrelation – we know that the SEs would change if these tests were conducted using models that account for autocorrelation._

    The order of the trend model was determined using "standard statistical procedures(e.g., _t_-test," with "either a linear or quadratic trend" used for each model.  The first problem with this approach is that the wording is imprecise: were t-tests the only basis for comparison?  Does the word "either" imply that the quadratic models did not contain a linear term (which they should)?  What were the p-values?
    The second problem with their method is that it was applied without any thought given to consistency.  The only species-variable combination with the same trend fit for each observation is $\delta_p$ (the isotropic composition, $\delta^{13}C ( \permil )$, of plant tissue) for Douglas Fir.  Unless there is a domain-knowledge-specific explanation  -- if there is one, it isn't given -- this practice seems questionable at best.

3) _Figure 3 contains two ACF plots. Discuss the patterns in the two SACFs (left two panels of Figure 3)._ 

4) _Note that their bounds for testing in Figure 3 are different from the formula provided in the book – we will discuss the modification I think they used later. What is the sample size (length of time series!) for the mean douglas-fir and mean western white pine time series. Based on these sample sizes, what cut-offs would you plot on Figure 3 based on the approximate result we discussed in class? How do theirs differ?_

They use +/- 2 SE for the bound

5) _We discussed four different reasons for doing time series modeling at the beginning of the semester (pages 1 and 2 of lecture notes). Of those four reasons, which one does their use of ARMA models fall into and why? There may be multiple choices that are reasonable here, so the argument you make is important. I am not talking about what you could do with their models, just what they did do with them and their reason for employing them._

I'd say some combination of description and accommodation

6) _Use arima.sim to simulate from Delta response AR(2) models for either the mean D or mean W discussed in Table 4 generating the same number of observations as in their time series (see Table 2). You can use a standard deviation for the white noise process of 0.4. This simulation is just concerned with the residual process after removing the trend. Plot a simulated fake time series, the SACF from the fake time series, and compare the results to those that you can find in the paper in Figure 3. How are your results similar or different from their results? Simulate a second series and discuss the differences in your two simulations._

```{r p6}

```

7) _Use R to find the roots of your selected AR(2) model based on their published coefficients and note what that tells you about the properties of the AR(2) process under consideration._

```{r p7}

```